{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-08-20 thesis/education.tex
discuss significance for submitted bug counts

{\bf Statistical significance.}
Table~\ref{tab:table-of-ze-studence} presents more detailed statistics from these six distributions.
To assess whether the differences are significant,
% discrete == I used "version 2" of the AD metric that R outputs.
I use the $k$-sample discrete Anderson-Darling test \cite{anderson-darling}
provided by the R statistical programming environment \cite{r-lang,r-ksamples},
comparing each control group to its respective Landslide users group.
Anderson-Darling uses a weighted sum-of-squares distance metric to compare two CDFs,
testing if their samples are likely to arise from the same underlying, though unspecified, distribution%
\footnote{
I choose Anderson-Darling over the simpler Kolmogorov-Smirnov,
which computes only the maximum instantaneous distance between CDFs,
for its better sensitivity both to repeated deviations and to tail differences \cite{beware-kolmogorov-smirnov}.
}.
I deem the difference in grade distributions significant when $p < 0.05$%
% FIXME: is this legit?
% should you be using the combined/grouped AD test instead?
\footnote{I choose not to correct for multiplicity among these 4 comparsons (see next section)
because each one is testing a different hypothesis.
},
although as in all $p$-value calculations,
this rejects only the null hypothesis,
not alternative hypotheses (such as ``more skilled students were more likely to sign up to begin with'').
%
Ultimately, only the same-semester comparisons were significant.
However, the distribution difference being smaller in P3 than in P2
% TODO: is there a good way to treat this more formally?
further suggests, albeit informally,
that the larger impact in P2 is not attributable {\em only} to selection bias.
% let me have this exclamation mark. please
That aside, ``Landslide helps students submit better P2s'' is the next most likely hypothesis I can think of!

@@

\begin{enumerate}
\item
{\bf Thread exit/join.}
%During {\tt thr\_exit()}, a thread must arrange for its own stack memory allocation to be freed.
After a thread finishes exiting, the memory allocated for its stack must be reclaimed for other uses.
A common student pitfall is to allow a {\tt thr\_join()}ing thread to free said stack space
while {\tt thr\_exit()} is still executing userspace C code (which typically accesses the stack),
or even for {\tt thr\_exit()} to free it itself.
In the former case, threads must interleave specifically to exhibit a use-after-free;
in the latter case, the use-after-free will be deterministic (i.e., present in all interleavings).
Other than Landslide, however,
the students have no Valgrind-like heap debugging tool which would report a bug immediately upon any illegal heap access.
This means that a subsequent {\tt thr\_create()} invocation would need to race to recycle the memory for a new thread stack
and conflict with the old exiting thread
before any problem could be detected.
% TODO: mark which of the tej bugs were deterministic vs race
The Landslide-friendly test {\tt thr\_exit\_join} is most likely to expose this bug.
%, and
%the official rubric penalties for it are
%{\tt :exit\_stack\_minor -2} and
%{\tt :exit\_stack\_race -4} (for the former case),
%and
%{\tt :uses\_freed\_stack -4} (for the latter).
Whether each student's submitted P2 solved or suffered from this bug
was determined by me personally inspecting their code.
%
This bug is especially remarkable in that not a single student who did not use Landslide these semesters
submitted a correct implementation.

@@

{\bf Statistical significance.}
% two-tailed btw, the better (more conservative) way,
% tbh one tailed might even make sense here, but it doesn't change anything wrt the cutoffs anyway, so
The $p$ values
in Table~\ref{tab:eval-common-bugs}
are calculated in R \cite{r-lang} using Fisher's exact test \cite{fishers-exact-test},
treating each row as an independent 2x2 contingency table.
I divide 0.05, the standard significance cutoff,
by the number of bugs to conservatively account for multiplicity \cite{xkcd-jellybeans}
(not that it matters).
% TODO write more

>>>>

{\bf Statistical significance.}
Table~\ref{tab:table-of-ze-studence} presents more detailed statistics from these six distributions.
To assess whether the differences are significant,
% discrete == I used "version 2" of the AD metric that R outputs.
I use the $k$-sample discrete Anderson-Darling test \cite{anderson-darling}
provided by the R statistical programming environment \cite{r-lang,r-ksamples},
comparing each control group to its respective Landslide users group.
Anderson-Darling uses a weighted sum-of-squares distance metric to compare two CDFs,
testing if their samples are likely to arise from the same underlying, though unspecified, distribution%
\footnote{
I choose Anderson-Darling over the simpler Kolmogorov-Smirnov,
which computes only the maximum instantaneous distance between CDFs,
for its better sensitivity both to repeated deviations and to tail differences \cite{beware-kolmogorov-smirnov}.
}.
I deem the difference in grade distributions significant when $p < 0.05$%
% FIXME: is this legit?
% should you be using the combined/grouped AD test instead?
\footnote{I choose not to correct for multiplicity among these 4 comparsons like I do in the next section
because each tests a different hypothesis.
},
although as in all $p$-value calculations,
this rejects only the null hypothesis,
not alternative hypotheses (such as ``more skilled students were more likely to sign up to begin with'').
%
Ultimately, only the same-semester comparisons were significant.
However, the distribution difference being smaller in P3 than in P2
% TODO: is there a good way to treat this more formally?
further suggests, albeit informally,
that the larger impact in P2 is not attributable {\em only} to selection bias.
% idk about this
%Though ``P3 grading criteria results in a tighter bell curve than P2'' is another potential hypothesis,
% let me have this exclamation mark. please
That aside, ``Landslide helps students submit better P2s'' is the next most likely hypothesis I can think of!

@@

\begin{enumerate}
\item
{\bf Thread exit/join.}
%During {\tt thr\_exit()}, a thread must arrange for its own stack memory allocation to be freed.
After a thread finishes exiting, the memory allocated for its stack must be reclaimed for other uses.
A common student pitfall is to allow a {\tt thr\_join()}ing thread to free said stack space
while {\tt thr\_exit()} is still executing userspace C code (which typically accesses the stack),
or even for {\tt thr\_exit()} to free it itself.
In the former case, threads must interleave specifically to exhibit a use-after-free;
in the latter case, the use-after-free will be deterministic (i.e., present in all interleavings).
Other than Landslide, however,
the students have no Valgrind-like heap debugging tool which would report a bug immediately upon any illegal heap access.
This means that a subsequent {\tt thr\_create()} invocation would need to race to recycle the memory for a new thread stack
and conflict with the old exiting thread
before any problem could be detected.
The Landslide-friendly test {\tt thr\_exit\_join} is most likely to expose this bug.
%, and
%the official rubric penalties for it are
%{\tt :exit\_stack\_minor -2} and
%{\tt :exit\_stack\_race -4} (for the former case),
%and
%{\tt :uses\_freed\_stack -4} (for the latter).
Whether each student's submitted P2 solved or suffered from this bug
was determined by me personally inspecting their code.
%

@@

{\bf Statistical significance.}
% two-tailed btw, the better (more conservative) way,
% tbh one tailed might even make sense here, but it doesn't change anything wrt the cutoffs anyway, so
The $p$ values
in Table~\ref{tab:eval-common-bugs}
are calculated in R \cite{r-lang} using Fisher's exact test \cite{fishers-exact-test},
treating each row as an independent 2x2 contingency table.
I divide 0.05, the standard significance cutoff,
by the number of bugs to conservatively account for multiplicity \cite{xkcd-jellybeans}
(not that it matters with these $p$ values).

The latter three bugs' occurences are thoroughly uncorrelated with Landslide use,
the {\tt thr\_exit()} bug standing alone with extremely high significance:
not a single student who did not use Landslide these semesters
submitted a correct implementation.
This difference is easily explainable:
the class-provided unit tests already do a good enough job catching the other three bugs
that students are able to find and fix them before submission regardless of what testing tool they used.
(The high correct-to-buggy ratio corroborates this.)
Nevertheless, that does not mean Landslide is pointless for these bugs;
it may well have helped the students reproduce them more reliably and/or diagnose them faster.
This is merely a null result for submission quality, not necessarily a negative one,
and the next section will attempt to evaluate such quality-of-life improvements instead.

On the other hand, because the {\tt thr\_exit()} bug typically manifests as a use-after-free,
the official tests must stress the thread library until the memory in question gets re-allocated
in just the right interleaving pattern to cause corruption that leads to a visible crash.
Landslide, on the other hand, detects this error immediately upon access (\sect{\ref{sec:landslide-valgrind-mode}}),
with no need for such complex corruption conditions.
Among the 11 Landslide users who submitted correct {\tt thr\_exit()}s in this regard,
it is difficult to say whether the heap checking alone, together with unit or stress tests,
would have been sufficient,
or whether heap checking and model checking were both necessary in concert to help them.
To address this doubt,
I simulated who among the buggy {\tt thr\_exit()} implementations
would have found and fixed the bug
given a heap-checking unit test
by checking whether Landslide finds their bug ``deterministically''
% TODO - more specific section reference - quicksand det/nondet drpps eval
(i.e., on the first interleaving, similarly to \sect{\ref{sec:quicksand-eval}}).
So doing,
the 11-10-0-16 distribution becomes
NN-MM-PP-QQ, % TODO
with a new $p$ value of
0.ppp, % TODO
%% which one, will it be??? :-O
% still a significant difference.
% still positively correlated, but no longer significant.
whose significance is a MYSTERYYYY.
